Lessons Learned for JWST Transiting Planet Time Series from Ground-based Studies of the NIRCam Detectors

### 2019-05-12 ms.tex
spelling out the read noise/shot noise calculations more clearly

It is instructive to look at dark frames to isolate the contributions of the read-out electronics, including the side-car asic, and how much they contribute to the error budget.
The dark frames do not contain the photon noise in the background nor the issues with the photometric stability of illumination lamps, which are a persistent challenge with ground-based tests.
We study here the contributions from read noise, 1/f noise and dark current.
However, there could be subtle changes in behavior of the electronic read noise under illumination, such as cross-talk where non-illuminated pixels in the images correlate with illuminated ones.

The standard long dark exposures used for detector characterization are a single integration that is 108 groups of 1 frame each (so-called RAPID mode).
The detectors are read in full frame mode so that each frame is 10.7 sec in duration with 4 output amplifiers in use.
These long dark integrations are used to create simulated time series as if it were observing many consecutive integrations in full frame mode.
The long darks are broken into 54 sequential pairs of images that are subtracted from each other (integration 1 minus integration 0, 3 minus 2, 5 minus 4 etc. in a zero-based counting scheme).
Figure \ref{fig:longDarkPhot} shows an example dark read pair from frame 55 minus frame 54.
In the raw data, there are 4 prominent vertical stripes that are 512 pixels wide and 2048 pixels tall due to offsets in the read-out amplifiers.
These amplifier offsets are caused by drifts in the reference voltage that are reset once per frame and vary from amplifier to amplifier.
Additionally, there are obvious correlations along the fast-read (X) direction due to 1/f noise that appear as horizontal stripes in the image.

As seen in Figure \ref{fig:longDarkPhot}, reference pixels can be used to reduce the large offsets between amplifiers.
The reference pixels (making up a 4 pixel boundary on the edges of the detector) are un-illuminated and can therefore be used to measure common-mode electronic noise without being affected by photon counting noise.
The reference pixels are most useful for subtracting the offsets between amplifiers because there are 8 rows of 512 pixels or a total of 2048 reference pixels to average for each amplifier.
However, the 1/f noise that correlates along the fast-read (X) direction is more difficult to remove with reference pixels.
Only 4 pixels on the left and 4 pixels on the right side may be used to subtract the 1/f noise.
Therefore, significantly residuals remain along the fast-read (X) direction in the reference-corrected read pair shown in Figure \ref{fig:longDarkPhot}.

@@

For a read pair subtraction, the ``CDS'' (correlated double sample) read noise is 13.3 e$^-$ or 7.2 DN for the A5 detector studied here.
If we assume each pixel's read noise is uncorrelated with its neighbors, the total noise within the source aperture is then $\sqrt{N_{px}}$ where $N_{px}$ is the number of pixels and $RN$ is the read noise of a single pixel.
If we also assume that the background pixels are uncorrelated with each other and the source pixels, then the total noise in background-subtracted photometry is 1309 DN.
As an order of magnitude estimate for photometric stability, this read noise should be compared with the 3$\times 10^{6}$ DN/s (or 3$\times 10^7$ DN for a single full frame) in our CV3 detector stability test, so the read noise is expected to contribute $\sim$ 34 ppm per frame.
When averaging together $\sim$ 100 full frames over $\sim$ 60 minutes, this would only contribute 3 ppm to the transit depth uncertainty.

>>>>

It is instructive to look at dark frames to isolate the contributions of the read-out electronics, including the side-car asic, and how much they contribute to the error budget.
The dark frames do not have issues with the photometric stability of illumination lamps, which is a persistent challenge with ground-based tests.
They also do not contain photon noise in the background, which can mask the electronic noise sources.
We study here the contributions from read noise, 1/f noise and dark current.
This experiment with dark frames does not test subtle changes in behavior of the electronic read noise under illumination, such as cross-talk where pixels that have no direct flux from a source in an image correlate with other regions of the detector with high flux rates.

The standard long dark exposures used for detector characterization are a single integration that is 108 groups of 1 frame each (so-called RAPID mode).
The detectors are read in full frame mode so that each frame is 10.7 sec in duration with 4 output amplifiers in use, thus the total time for the exposure was 19.3 min.
These long dark integrations are used to create simulated time series as if it were observing many consecutive integrations in full frame mode.
The long darks are broken into 54 sequential pairs of images that are subtracted from each other (integration 1 minus integration 0, 3 minus 2, 5 minus 4 etc. in a zero-based counting scheme).
Figure \ref{fig:longDarkPhot} shows an example dark read pair from frame 55 minus frame 54.
In the raw data, there are 4 prominent vertical stripes that are 512 pixels wide and 2048 pixels tall due to offsets in the read-out amplifiers.
These amplifier offsets are caused by drifts in the reference voltage that are reset once per frame and vary from amplifier to amplifier.
Additionally, there are obvious correlations along the fast-read (X) direction due to 1/f noise that appear as horizontal stripes in the image.

As seen in Figure \ref{fig:longDarkPhot}, reference pixels can be used to reduce the large offsets between amplifiers.
The reference pixels (making up a 4 pixel boundary on the edges of the detector) are un-illuminated and can therefore be used to measure common-mode electronic noise without being affected by photon counting noise.
The reference pixels are most useful for subtracting the offsets between amplifiers because there are 8 rows of 512 pixels or a total of 4096 reference pixels to average for each amplifier at the bottom and top of the arrays.
There are an additional 4 columns on the left and right side boundaries (8176 pixels) that can be used to further correct the left and right amplifier offests.
However, the 1/f noise that correlates along the fast-read (X) direction is more difficult to remove with reference pixels.
Only 4 pixels on the left and 4 pixels on the right side may be used to subtract the 1/f noise.
Therefore, significantly residuals remain along the fast-read (X) direction in the reference-corrected read pair shown in Figure \ref{fig:longDarkPhot}.
These side reference pixels also cannot give any information about the read noise correlations that happen on shorter length scales than 512 pixels (5 ms).

@@

For a read pair subtraction, the ``CDS'' (correlated double sample) read noise is 13.3 e$^-$ or 7.2 DN for the A5 detector studied here.
If we assume each pixel's read noise is uncorrelated with its neighbors, the total noise within the source aperture is then $\sqrt{N_{px}} \times RN $ where $N_{px}$ is the number of pixels and $RN$ is the read noise of a single pixel.
If we also assume that the background pixels are uncorrelated with each other and the source pixels, then the total noise in background-subtracted photometry is 1309 DN.

This theoretical read noise should be compared to the photon noise expected inside an aperture during photometric stability time series applications.
The number of DN collected within a 70 pixel radius aperture for WLP8 data in our CV3 detector stability test with 2 groups in RAPID (ie a correlated double sample) was $DN_{tot}=3.9\times 10^7$ DN during the integration.
The peak pixel counts in this integration were 30,000 DN or about 60\% well depth for the A1 detector.
The Poisson shot noise for photons is $N_{phot} = \sqrt{DN_{tot}} / \sqrt{G} = 4,400$DN, where $G$ is the gain ($G \approx 2.0$ for the SW detectors).
In terms of fractional noise, the photon shot noise should be 113 ppm and the read noise should be 33 ppm.
Therefore, the read noise is expected to be 1/4 the shot noise for 2 group RAPID read mode (ie CDS).
When averaging together $\sim$ 100 full frames over $\sim$ 60 minutes, this would only contribute 3 ppm to the transit depth uncertainty.
